### 2021-04-27 notes/fischer.tex
Finsihed the proof Fischer's inequality.

\begin{document}
\noindent{\Huge \textbf\textsf{{The generalized determinant and the Fischer's inequality}}}

@@

\begin{lemma}
	We have the following properties (again, $r$ is the rank of $A$):
	\begin{enumerate}
		\item For $\rho\ne 0$ we have $\detp (\rho A) = \rho^{n-r} \detp (A)$. 
		\item For $A\in M_n(\C)$ and $B\in M_m(\C)$ we have
		\begin{equation}
			\detp \begin{pmatrix}
				A & 0 \\
				0 & B 
			\end{pmatrix} = \detp A\; \detp B.
		\end{equation}
		\item For any unitary $U$ we have 
		\begin{equation}
			\detp U A U^\star = \detp A.
		\end{equation} 
	\end{enumerate}
\end{lemma}

@@

\begin{lemma}
	Every positive semidefinite matrix $A$ has the row and column inclusion property.
\end{lemma}

@@

>>>>

\begin{document}
\noindent{\Huge \textbf\textsf{{The generalized determinant and  Fischer's inequality}}}

@@

\begin{lemma}
	We have the following properties (again, $r$ is the rank of $A$):
	\begin{enumerate}
		\item For $\rho\ne 0$ we have $\detp (\rho A) = \rho^{n-r} \detp (A)$. 
		\item For $A\in M_n(\C)$ and $B\in M_m(\C)$ we have
		\begin{equation}
			\detp \begin{pmatrix}
				A & 0 \\
				0 & B 
			\end{pmatrix} = \detp A\; \detp B.
		\end{equation}
		\item For any unitary $U$ we have 
		\begin{equation}
			\detp U A U^\star = \detp A.
		\end{equation}
		\item If $\mathcal{N}(A)=\{0\}$, i.e. if $\rank(A)=n$ we have
		\begin{equation}
			\detp A = \det A.
		\end{equation} 
	\end{enumerate}
\end{lemma}

@@

\begin{lemma}\label{lem.columnincl}
	Every positive semidefinite matrix $A$ has the row and column inclusion property.
\end{lemma}

@@

\begin{remark}
	A simple example illustrates the importance of the condition on the ranks. Let
	\begin{equation}
		H = \begin{pmatrix}
			1 & a \\
			a & a^2
		\end{pmatrix}.
	\end{equation}
	Then $\det H = 0$ and $\detp H = 1 + a^2$ (the two eigenvalues of $H$ are $0$ and $1+a^2$ with eigenvectors $(a, -1)^T$ and $(1, a)^T$ respectively). For $a>0$ we have 
	\begin{equation}
		\detp 1\detp a^2 = a^2.
	\end{equation}
	$H$ is Hermitian and positive semidefinite but Fischer's inequality does not hold for the generalized determinant. Note that 
	\begin{equation}
		\rank(H) = 1,
	\end{equation}
	while $\rank(1) + \rank(a^2) = 2$. 
\end{remark}

\section{Improvement}
We now want ti improve the theorem by dealing with case of equality:

\begin{theorem}
	Let $H$ be Hermitian and positive semidefinite and let $H$ be partitioned as follows:
		\begin{equation}
		H = \begin{pmatrix}
			A & B \\
			B^\star & C
		\end{pmatrix},
	\end{equation}
	with $A\in M_k(\C)$ and $1\le k< n$. Assume that 
	\begin{equation}\label{eqn.rankadd}
		\rank(H) = \rank(A) + \rank(C), 
	\end{equation}
	then
	\begin{equation}
		\detp H \le \detp A\; \detp C,
	\end{equation}
	with equality if and only if $B=0$.
\end{theorem}

\begin{proof}
	Since $A$ and $C$ are Hermitian there are unitary matrices $U=U^\star\in M_k(\C)$ and $V=V^\star\in M_{n-k}(\C)$ such that
	\begin{align}
		U A U^\star & = \begin{pmatrix}
			\lambda_1 & 0 & 0 \\
			0 & \ddots & 0 \\
			0 & 0 & \lambda_k
		\end{pmatrix} \\
		& \nonumber \\
		V C V^\star & = \begin{pmatrix}
			\mu_1 & 0 & 0 \\
			0 & \ddots & 0 \\
			0 & 0 & \mu_{n-k}
		\end{pmatrix}.
	\end{align}
	With
	\begin{equation}
		W = \begin{pmatrix}
			U & 0 \\
			0 & V
		\end{pmatrix}
	\end{equation}
	we have
	\begin{equation}
		W H W^\star = \begin{pmatrix}
 	\begin{matrix}
			\lambda_1 &  &  \\
			 & \ddots &  \\
			 &  & \lambda_k
		\end{matrix} & UBV^\star \\
		VB^\star U^\star & \begin{matrix}
			\mu_1 &  &  \\
			 & \ddots &  \\
			 &  & \mu_{n-k}
		\end{matrix}
 	\end{pmatrix}
	\end{equation}
	We now make use of the fact that $W H W^\star$ is positive semidefinite (since $H$ is positive semidefinite). Let $r_A$ be the rank of $A$ and $r_C$ be the rank of $C$. We assume that we have ordered the eigenvalues in such a way that 
	\begin{equation}
		\lambda_i = 0, \text{\ for\ } i>r_A,
	\end{equation} 
	and
		\begin{equation}
		\mu_j = 0, \text{\ for\ } j>r_C.
	\end{equation}
	Because $W H W^\star$ is positive semidefinite it follows that
	\begin{equation}\label{eqn.lotsofzeroes}
		(UBV^\star)_{i,j}=0, \text{\ for\ } i>r_A \text{\ or\ }j>r_C.
	\end{equation}
	Note the condition here. It is sufficient that either $i>r_A$ \emph{or} $j>r_C$ for $(UBV^\star)_{i,j}$ to vanish (this follows from lemma \ref{lem.columnincl}).

	We now concentrate on the elements of $W H W^\star$ that are not zero. Let us define the following set of indices 
	\begin{equation}
		\iota = \{1,\ldots,r_A,k+1,\ldots,k+r_C\},
	\end{equation}
	and let us set $G = W H W^\star$. We have just established that
	\begin{equation}
		G[\iota^c] = 0.
	\end{equation}
	The remaining nonzero components are contained in $G[\iota]$. Because we are assuming that 
	\begin{equation}
		\rank(H) = \rank(A) + \rank(C),		
	\end{equation}
	and since 
	\begin{equation}
		\rank(G[\iota]) = \rank(H),	
	\end{equation}
	 $G[\iota]$ is a Hermitian positive definite matrix. Furthermore we have
	 \begin{align}
	 	\detp H & = \det G[\iota] \\
	 	& \le \lambda_1\cdots\lambda_{r_A}\cdot\mu_1\cdots\mu_{r_C} \\
	 	& = \detp A\; \detp C,
	 \end{align}
	 with equality if and only if the rest of $UBV^\star$ is zero. \qed
\end{proof}

\begin{remark}
	Equation (\ref{eqn.lotsofzeroes}) is equivalent to $\mathcal{N}(C)\subset\mathcal{N}(B)$ and $\mathcal{N}(A)\subset\mathcal{N}(B^\star)$. These inclusions were established in lemma \ref{lem.columnincl}. 	
\end{remark}
